### 2024-01-02 Chapter5.tex
Added gelatin HSI forwards to C5

\section{Introduction}
As described in Chapter \ref{chap:intro}, hyperspectral imaging (HSI) is able to capture spatially resolved, diffuse reflectance spectra via non-contact imaging. It has been posed that this could enable quantitative, spatially-resolved, tissue oxygenation ($StO_2$) extraction from intra-operative HSI data, with the analytical models investigated in Chapters \ref{chap:1layer} and \ref{chap:2layer} being some of the prominent methods of doing so.

HSI data is often of lower spectral resolution than the spectrophotometer data tested in the previous chapters, particularly that of snapshot HSI data which compromises spectral resolution in favour of shorter aquisition times. Additionally, these sensors may have increased noise levels due to the lack of spatial averaging.

In this chapter, I will quantify the impact of reducing the spectral resolution and introduction of noise on the performance of the analytical models investigated in Chapters \ref{chap:1layer} and \ref{chap:2layer} by simulating a snapshot camera response from the Monte Carlo datasets used in the previous chapters. These impacts are then evaluated with experimental HSI data taken of the phantoms detailed in Chapter \ref{chap:1layer}. Finally, the models are applied to a subset of neurosurgical HSI data obtained intra-operatively as part of the HELICoiD and NHSI datasets.

@@

\subsection{Hyperspectral imaging of gelatin-based phantoms}\label{sec:imagingphantoms}
To obtain a measured hyperspectral dataset with known ground truth, HSI images are taken of the two-dye configuration gelatin-based phantoms detailed in Section \ref{sec:methodsphantoms}. The imaging configuration is similar to that described in Section \ref{materials} using a 4x4 16 band visible range snapshot mosaic camera (Ximea utilising the IMEC CMV2K-SSM4X4-VIS sensor, Germany) alongside an f=35mm coupler (Karl Storz, NDTec, VCam HD-F-35 - Camera Lens Adapter, Germany) and a $0^\circ$ exoscope (Karl Storz Endoscopy, VITOM Telescope 0° w Integ. Illuminator, UK), as detailed in \cite{Ebner2021} and pictured in Figure \ref{fig:imagephantoms}.
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.3\textwidth]{imagephantoms.jpg}
    \caption{Depiction of hyperspectral imaging set-up used to acquire images of gelatin-based tissue phantoms.}
    \label{fig:imagephantoms}
\end{figure}
The configurations imaged are detailed in Table \ref{tb:imagedphantoms} and cover the majority of the 2-dye configuration dataset that remained viable after two weeks of refrigeration. 
\begin{table}[ht!]
    \centering
    \caption{Table displaying the ratios of dyes [acid red 1 (AR1), acid red 14 (AR14), crystal violet (CV)] a used for each dye configuration alongside the total dye scaled concentration in arbitrary units. The intralipid (v/v \%) imaged for each configuration are also indicated.}
    \begin{tabular}{|c|c|c|c|c|c|c|c|}
        \hline
        Total dye scaled & \multicolumn{2}{|c|}{Dye } & \multicolumn{5}{c|}{Intralipid } \\
        concentration & \multicolumn{2}{|c|}{ratio} & \multicolumn{5}{c|}{concentrations (v/v \%)} \\
        (arbitrary units) & AR1 & AR14 & 1.0 & 2.25 & 3.5 & 4.75 & 6 \\
        \hline
        1 & 1 & 1 & & x & x & x & \\
        10 & 1 & 0 & x & x & x & x & x \\
        10 & 3 & 1 & x & x & x & x & x \\
        10 & 1 & 1 &  & x & x & x & x \\
        10 & 1 & 3 & x & x & x & x & x \\
        10 & 0 & 1 & x & x &  & x & \\
        20 & 1 & 1 &  & & x & x &  \\
        \hline
    \end{tabular}
    \label{tb:imagedphantoms}
\end{table}
Each image is annotated using ImFusion Labels software to capture pixels corresponding to phantoms while avoiding regions with high specular reflection.

\subsection{Neurosurgical HSI datasets}\label{sec:NeuroHSIdata}
Two neurosurgical HSI datasets are leveraged in this preliminary work. Initially, the HELICoiD dataset is used which captures data in the wavelength range 400-1000nm using a pushbroom spatial scanning HSI camera\cite{Fabelo2019}. This provides 36 images of high spectral resolution which are annotated with use of spectral angle mapper to identify four key classes: "normal tissue, tumor tissue, blood vessel, and background elements". For this work we consider only the spectral range of 450-600nm as this is the range of use for the analytical models considered. Secondly, a NeuroHSI dataset is considered which is obtained using the same snapshot mosaic camera (Ximea utilising the IMEC CMV2K-SSM4X4-VIS sensor, Germany) as in Section \ref{sec:imagingphantoms} whose bands are within the 450-600nm wavelength range. In contrast to Section \ref{sec:imagingphantoms}, this is mounted to a neurosurgical microscope (Zeiss Kinevo 900, Germany) for simpler integration into the neurosurgical workflow, allowing for use in a wide variety of neurosurgical cases. This dataset is manually annotated using ImFusion Labels software to identify over 30 classes.

@@

\subsection{Monte Carlo}
The forwards analytical models are assessed using simulated camera responses generated from Monte Carlo data. 
%The Yudovsky 2009 and Jacques 1999 single-layer models are evaluated for three refractive indices, whereas the Yudovsky 2009 double-layer model is evaluated for $n=1.44$ only as in Chapters \ref{chap:1layer} and \ref{chap:2layer}. 
These are evaluated in Table \ref{tb:forwardsHSIMC} for $n=1.44$ (further refractive indices can be found in Appendices \ref{ap:forwardsHSIMCq} and \ref{ap:forwardsHSIMCr}) and examples for the single and double layer configurations shown in Figure \ref{fig:forwardsHSIMC}. 
\begin{table}[bhp] %could remove noise levels here
    \centering
    \caption{Mean (standard deviation) $NRMSE$ (3.d.p.) between the simulated camera responses of each forwards spectrum from each model and each of 100 Monte Carlo simulated spectra using the same ground truth variable parameters for quantitative and relative data at a variety of noise levels (0, 0.01, 0.03). This is presented with the Pearson $r$ (bold if Pearson $p < 0.05$) for the linear regression between all forwards spectra against Monte Carlo simulated spectra for each refractive index dataset and each analytical model. All metrics are evaluated for the wavelength region of 450-600nm.}
    \begin{tabular}{|cc|c|ccc|ccc|}
        \hline
        Model & Layers & Quantitative & \multicolumn{3}{|c}{$NRMSE$} & \multicolumn{3}{|c|}{$r$} \\
         & & (Q) or & 0 & 0.01 & 0.03 & 0 & 0.01 & 0.03 \\
         & & Relative (R) &  &  &  &  & &  \\
        \hline
        \multirow{2}{*}{\shortstack{Yudovsky\\ 2009}} & \multirow{2}{*}{1} & Q & \shortstack{0.009 \\(0.004)} & \shortstack{0.082 \\(0.057)} & \shortstack{0.224 \\(0.125)} & \textbf{1.000} & \textbf{0.993} & \textbf{0.942} \\
        & & R & \shortstack{0.003 \\(0.002)} & \shortstack{0.075 \\(0.044)} & \shortstack{0.212 \\(0.118)} & \textbf{1.000} & \textbf{0.916} & \textbf{0.602} \\
        \hline
        \multirow{3}{*}{\shortstack{Jacques\\ 1999}} & \multirow{2}{*}{1} & Q & \shortstack{0.026 \\(0.046)} & \shortstack{0.089 \\(0.077)} & \shortstack{0.227 \\(0.130)} & \textbf{1.000} & \textbf{0.993} & \textbf{0.942} \\
        & & R & \shortstack{0.014 \\(0.021)} & \shortstack{0.076 \\(0.045)} & \shortstack{0.212 \\(0.117)} & \textbf{0.995} & \textbf{0.915} & \textbf{0.605} \\
        \hline
        \multirow{2}{*}{\shortstack{Yudovsky\\ 2009}} & \multirow{2}{*}{2} & Q & \shortstack{0.166 \\(0.133)} & \shortstack{0.482 \\(0.289)} & \shortstack{0.738 \\(0.294)} & \textbf{0.992} & \textbf{0.971} & \textbf{0.829} \\
         &  & R & \shortstack{0.033 \\(0.026)} & \shortstack{0.428 \\(0.286)} & \shortstack{0.684 \\(0.286)} & \textbf{0.991} & \textbf{0.270} & -0.007 \\
        \hline
    \end{tabular}
    \label{tb:forwardsHSIMC}
\end{table}

\begin{figure}[htbp]
    \centering
    \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{forwards_noise_Q.pdf}
        \caption{}
        \label{fig:egforwardsnoiseQ}
    \end{subfigure}
    \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{forwards_noise_R.pdf}
        \caption{}
        \label{fig:egforwardsnoiseR}
    \end{subfigure}
    \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{forwards_noise2_Q.pdf}
        \caption{}
        \label{fig:egforwards2noiseQ}
    \end{subfigure}
    \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{forwards_noise2_R.pdf}
        \caption{}
        \label{fig:egforwards2noiseR}
    \end{subfigure}
    \caption{Top: Example of the simulated camera response spectra from each forward single-layer analytical model: Yudovsky 2009 (\textcolor{MyGreen}{green dotted}) and Jacques 1999 (\textcolor{MyOrange}{orange dashed}), using ground truth variables for a refractive index of 1.44 compared to that predicted by Monte Carlo with (\textcolor{purple}{purple dash-dotted} and without noise added (\textcolor{MyBlue}{blue solid}) for quantitative (left) and relative (right) data. Bottom: Demonstration of the variability of quality of prediction between the simulated camera responses from the Yudovsky 2009 two layer model (\textcolor{MyGreen}{green dotted}) and Monte Carlo with (\textcolor{purple}{purple dash-dotted} and without noise added (\textcolor{MyBlue}{blue solid}) when using the same ground truth parameters for quantitative (left) or relative (right) data.}
    \label{fig:forwardsHSIMC}
\end{figure}

@@

\begin{table}[htb!]
    \centering
    \caption{The Pearson $r$ values (bold if $p<0.05$) of the linear regression line between the fitted tissue parameters and their ground truth displayed with their median absolute percentage errors ($APE$). This is shown for each variable when extracted by fitting Yudovsky 2009 single layer (Y1), Jacques 1999 (J), or Yudovsky 2009 double layer (Y2) to the quantitative (Q) or relative (R) Monte-Carlo datasets for $n=1.44$ at each noise level (0, 0.01, 0.03). All presented to 3s.f.}
    \begin{tabular}{|ccc|ccc|ccc|}
        \hline
        Parameter & Model & Quantitative & \multicolumn{3}{c}{$r$} & \multicolumn{3}{|c|}{median} \\
        & & (Q) or & \multicolumn{3}{c}{} & \multicolumn{3}{|c|}{$APE$ (\%)} \\
        & & Relative (R) & 0 & 0.01 & 0.03 & 0 & 0.01 & 0.03 \\
        \hline
        \multirow{6}{*}{$StO_2$} & \multirow{2}{*}{Y1} & Q & \textbf{0.999} & \textbf{0.825} & \textbf{0.369} & 1.90 & 24.8 & 85.2 \\
        & & R & \textbf{0.999} & \textbf{0.752} & \textbf{0.410} & 2.23 & 41.4 & 76.7 \\
        \cline{2-9}
        & \multirow{2}{*}{J} & Q & \textbf{0.978} & \textbf{0.829} & \textbf{0.402} & 4.06 & 22.5 & 83.6 \\
        & & R & \textbf{0.977} & \textbf{0.755} & \textbf{0.447} & 4.77 & 40.3 & 76.7 \\
        \cline{2-9}
        & \multirow{2}{*}{Y2} & Q & \textbf{0.751} & 0.0612 & 0.0194 & 16.8 & 66.5 & 68.9 \\
        & & R & \textbf{0.679} & \textbf{0.198} & 0.0448 & 18.8 & 63.0 & 75.1 \\
        \hline
        \multirow{6}{*}{$f_{blood}$} & \multirow{2}{*}{Y1} & Q & \textbf{0.984} & \textbf{0.667} & \textbf{0.450} & 5.83 & 23.7 & 51.8\\
        & & R & \textbf{0.870} & \textbf{0.547} & \textbf{0.535} & 16.7 & 47.5 & 51.1\\
        \cline{2-9}
        & \multirow{2}{*}{J} & Q & \textbf{0.924} & \textbf{0.719} & \textbf{0.498} & 7.14 & 25.0 & 46.7 \\
        & & R & \textbf{0.859} & \textbf{0.681} & \textbf{0.644} & 24.1 & 39.6 & 45.0 \\
        \cline{2-9}
        & \multirow{2}{*}{Y2} & Q & \textbf{0.872} & 0.147 & 0.0329 & 23.1 & 115 & 160 \\
        & & R & \textbf{0.609} & 0.189 & 0.0466 & 37.6 & 97.7 & 108 \\
        \hline
        \multirow{6}{*}{$a$} & \multirow{2}{*}{Y1} & Q & \textbf{0.992} & \textbf{0.773} & \textbf{0.558} & 3.85 & 18.7 & 35.1 \\
        & & R & \textbf{0.703} & 0.160 & \textbf{0.558} & 26.2 & 62.3 & 35.1 \\
        \cline{2-9}
        & \multirow{2}{*}{J} & Q & \textbf{0.961} & \textbf{0.759} & \textbf{0.542} & 4.69 & 17.1 & 30.4 \\
        & & R & \textbf{0.431} & -0.00356 & 0.166 & 27.7 & 57.7 & 51.7 \\
        \cline{2-9}
        & \multirow{2}{*}{Y2} & Q & \textbf{0.674} & 0.00365 & 0.0666 & 17.9 & 37.8 & 37.4 \\
        & & R & \textbf{0.333} & 0.0199 & -0.100 & 27.0 & 36.7 & 42.2 \\
        \hline
        \multirow{6}{*}{$b$} & \multirow{2}{*}{Y1} & Q & \textbf{0.998} & \textbf{0.751} & \textbf{0.516} & 2.63 & 21.8 & 52.0 \\
        & & R & \textbf{0.996} & \textbf{0.607} & \textbf{0.426} & 3.31 & 39.5 & 51.2 \\
        \cline{2-9}
        & \multirow{2}{*}{J} & Q & \textbf{0.934} & \textbf{0.757} & \textbf{0.515} & 5.70 & 23.7 & 50.4 \\
        & & R & \textbf{0.938} & \textbf{0.641} & \textbf{0.441} & 7.66 & 32.6 & 46.4 \\
        \cline{2-9}
        & \multirow{2}{*}{Y2} & Q & \textbf{0.735} & 0.169 & -0.00249 & 16.8 & 56.4 & 61.3 \\
        & & R & \textbf{0.360} & 0.0441 & -0.0711 & 33.2 & 61.7 & 63.6 \\
        \hline
        \multirow{2}{*}{$f_{mel}$} & \multirow{2}{*}{Y2} & Q & \textbf{0.852} & \textbf{0.622} & \textbf{0.477} & 28.0 & 85.1 & 90.1 \\ 
        & & R & \textbf{0.517} & 0.166 & \textbf{0.201} & 35.3 & 108 & 132 \\
        \hline
        \multirow{2}{*}{$L_1$} & \multirow{2}{*}{Y2} & Q & \textbf{0.824} & \textbf{0.401} & 0.123 & 36.7 & 69.5 & 84.5 \\
        & & R & \textbf{0.806} & 0.0777 & -0.0675 & 38.1 & 65.4 & 76.6 \\
        \hline
    \end{tabular}    
    \label{tb:backwardsHSIMC}
\end{table}

>>>>

\section{Introduction}
As described in Chapter \ref{chap:intro}, hyperspectral imaging (HSI) is able to capture spatially resolved, diffuse reflectance spectra via non-contact imaging\cite{Lu2014,Giannoni2018,Calin2014,Shapey2019}. It has been posed that this could enable quantitative, spatially-resolved, tissue oxygenation ($StO_2$) extraction from intra-operative HSI data, with the analytical models investigated in Chapters \ref{chap:1layer} and \ref{chap:2layer} being some of the prominent methods of doing so \cite{Yudovsky2009, Jacques1999, Clancy2015, Clancy2020}.

HSI data is often of lower spectral resolution than the spectrophotometer data tested in the previous chapters, particularly that of snapshot HSI data which compromises spectral resolution in favour of shorter aquisition times\cite{Geelen2014}. Additionally, Hyperspectral imaging is known to suffer from spectral noise which may be reduced significantly by spatial averaging until comparable to traditional optical methods \cite{Zhang2020}.

In this chapter, I will quantify the impact of reducing the spectral resolution and introduction of noise on the performance of the analytical models investigated in Chapters \ref{chap:1layer} and \ref{chap:2layer} by simulating a snapshot camera response from the Monte Carlo datasets used in these previous chapters. These impacts are then evaluated with experimental HSI data of the phantoms detailed in Chapter \ref{chap:1layer}. Finally, the models are applied to a subset of neurosurgical HSI data obtained intra-operatively as part of the HELICoiD and NHSI datasets.

@@

\subsection{Hyperspectral imaging of gelatin-based phantoms}\label{sec:imagingphantoms}
To obtain a measured hyperspectral dataset with known ground truth, HSI images are taken of the two-dye configuration gelatin-based phantoms detailed in Section \ref{sec:methodsphantoms}. The imaging configuration is similar to that described in Section \ref{materials} using a 4x4 16 band visible range snapshot mosaic camera (Ximea utilising the IMEC CMV2K-SSM4X4-VIS sensor, Germany) alongside an f=35mm coupler (Karl Storz, NDTec, VCam HD-F-35 - Camera Lens Adapter, Germany), a $0^\circ$ exoscope (Karl Storz Endoscopy, VITOM Telescope 0° w Integ. Illuminator, UK) and a Xenon light source (Karl Storz Endoscopy, Cold Light Fountain D light C, UK), as detailed in \cite{Ebner2021} and pictured in Figure \ref{fig:imagephantoms}.
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.3\textwidth]{imagephantoms.jpg}
    \caption{Depiction of hyperspectral imaging set-up used to acquire images of gelatin-based tissue phantoms.}
    \label{fig:imagephantoms}
\end{figure}
The configurations imaged are detailed in Table \ref{tb:imagedphantoms} and cover the majority of the 2-dye configuration dataset that remained viable after two weeks of refrigeration. 
\begin{table}[ht!]
    \centering
    \caption{Table displaying the ratios of dyes [acid red 1 (AR1), acid red 14 (AR14), crystal violet (CV)] a used for each dye configuration alongside the total dye scaled concentration in arbitrary units. The intralipid (v/v \%) imaged for each configuration are also indicated.}
    \begin{tabular}{|c|c|c|c|c|c|c|c|}
        \hline
        Total dye scaled & \multicolumn{2}{|c|}{Dye } & \multicolumn{5}{c|}{Intralipid } \\
        concentration & \multicolumn{2}{|c|}{ratio} & \multicolumn{5}{c|}{concentrations (v/v \%)} \\
        (arbitrary units) & AR1 & AR14 & 1.0 & 2.25 & 3.5 & 4.75 & 6 \\
        \hline
        1 & 1 & 1 & & x & x & x & \\
        10 & 1 & 0 & x & x & x & x & x \\
        10 & 3 & 1 & x & x & x & x & x \\
        10 & 1 & 1 &  & x & x & x & x \\
        10 & 1 & 3 & x & x & x & x & x \\
        10 & 0 & 1 & x & x &  & x & \\
        20 & 1 & 1 &  & & x & x &  \\
        \hline
    \end{tabular}
    \label{tb:imagedphantoms}
\end{table}
A white reference of a Spectralon tile of 95\% is taken under the same lighting conditions and imaging plane as the phantoms. This white reference is either used directly for white balancing quantitative data or to generate spectral sensitivities used to generate mean normalised relative as in Chapter \ref{chap:SWB}. This demonstrates both the highest fidelity and more realistic data processing pathways. Following this, bilinear demosaicing is used and cross-talk correction is applied per pixel of the image also detailed in Chapter \ref{chap:SWB}. 
Each image is manually annotated using ImFusion Labels software to capture pixels corresponding to phantoms while avoiding regions with high specular reflection. This allows a mean spectrum of the phantom to be generated and compared per pixel processing of the annotated region.

\subsection{Neurosurgical HSI datasets}\label{sec:NeuroHSIdata}
Two neurosurgical HSI datasets are leveraged in this preliminary work. Initially, the HELICoiD dataset is used which captures data in the wavelength range 400-1000nm using a pushbroom spatial scanning HSI camera\cite{Fabelo2019}. This provides 36 images of high spectral resolution which are annotated with use of spectral angle mapper to identify four key classes: "normal tissue, tumor tissue, blood vessel, and background elements". For this work we consider only the spectral range of 450-600nm as this is the range of use for the analytical models considered. Secondly, a NeuroHSI dataset is considered which is obtained using the same snapshot mosaic camera (Ximea utilising the IMEC CMV2K-SSM4X4-VIS sensor, Germany) as in Section \ref{sec:imagingphantoms} whose bands are within the 450-600nm wavelength range. In contrast to Section \ref{sec:imagingphantoms}, this is mounted to a neurosurgical microscope (Zeiss Kinevo 900, Germany) whose integrated light source and lenses are used for simpler integration into the neurosurgical workflow, allowing for use in a wide variety of neurosurgical cases. For each imaging timepoint, the position of the neurosurgical microscope is recorded and used to obtain a flatfield reference outside the sterile field in a closely correlated position and lighting conditions. Similarly to the phantom HSI processing pipeline, this is either used to provide quantitative data from flatfield imaging or used to generate spectral senstivities for relative data before demosaicing and cross-talk correction. These spectral sensitivities are found to be similar to those pre-calculated from the light source spectrum. This dataset is manually annotated using ImFusion Labels software to identify over 30 classes.

@@

\subsection{Monte Carlo}
The forwards analytical models are assessed using simulated camera responses generated from Monte Carlo data. 
%The Yudovsky 2009 and Jacques 1999 single-layer models are evaluated for three refractive indices, whereas the Yudovsky 2009 double-layer model is evaluated for $n=1.44$ only as in Chapters \ref{chap:1layer} and \ref{chap:2layer}. 
These are evaluated in Table \ref{tb:forwardsHSIMC} for $n=1.44$ (further refractive indices can be found in Appendices \ref{ap:forwardsHSIMCq} and \ref{ap:forwardsHSIMCr}) and examples for the single and double layer configurations shown in Figure \ref{fig:forwardsHSIMC}. 
\begin{table}[h!] %could remove noise levels here
    \centering
    \caption{Mean (standard deviation) $NRMSE$ (3.d.p.) between the simulated camera responses of each forwards spectrum from each model and each of 100 Monte Carlo simulated spectra using the same ground truth variable parameters for quantitative and relative data at a variety of noise levels (0, 0.01, 0.03). This is presented with the Pearson $r$ (bold if Pearson $p < 0.05$) for the linear regression between all forwards spectra against Monte Carlo simulated spectra for each refractive index dataset and each analytical model. All metrics are evaluated for the wavelength region of 450-600nm.}
    \begin{tabular}{|cc|c|ccc|ccc|}
        \hline
        Model & Layers & Quantitative & \multicolumn{3}{|c}{$NRMSE$} & \multicolumn{3}{|c|}{$r$} \\
         & & (Q) or & 0 & 0.01 & 0.03 & 0 & 0.01 & 0.03 \\
         & & Relative (R) &  &  &  &  & &  \\
        \hline
        \multirow{2}{*}{\shortstack{Yudovsky\\ 2009}} & \multirow{2}{*}{1} & Q & \shortstack{0.009 \\(0.004)} & \shortstack{0.082 \\(0.057)} & \shortstack{0.224 \\(0.125)} & \textbf{1.000} & \textbf{0.993} & \textbf{0.942} \\
        & & R & \shortstack{0.003 \\(0.002)} & \shortstack{0.075 \\(0.044)} & \shortstack{0.212 \\(0.118)} & \textbf{1.000} & \textbf{0.916} & \textbf{0.602} \\
        \hline
        \multirow{3}{*}{\shortstack{Jacques\\ 1999}} & \multirow{2}{*}{1} & Q & \shortstack{0.026 \\(0.046)} & \shortstack{0.089 \\(0.077)} & \shortstack{0.227 \\(0.130)} & \textbf{1.000} & \textbf{0.993} & \textbf{0.942} \\
        & & R & \shortstack{0.014 \\(0.021)} & \shortstack{0.076 \\(0.045)} & \shortstack{0.212 \\(0.117)} & \textbf{0.995} & \textbf{0.915} & \textbf{0.605} \\
        \hline
        \multirow{2}{*}{\shortstack{Yudovsky\\ 2009}} & \multirow{2}{*}{2} & Q & \shortstack{0.166 \\(0.133)} & \shortstack{0.482 \\(0.289)} & \shortstack{0.738 \\(0.294)} & \textbf{0.992} & \textbf{0.971} & \textbf{0.829} \\
         &  & R & \shortstack{0.033 \\(0.026)} & \shortstack{0.428 \\(0.286)} & \shortstack{0.684 \\(0.286)} & \textbf{0.991} & \textbf{0.270} & -0.007 \\
        \hline
    \end{tabular}
    \label{tb:forwardsHSIMC}
\end{table}

\begin{figure}[h!]
    \centering
    \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{forwards_noise_Q.pdf}
        \caption{}
        \label{fig:egforwardsnoiseQ}
    \end{subfigure}
    \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{forwards_noise_R.pdf}
        \caption{}
        \label{fig:egforwardsnoiseR}
    \end{subfigure}
    \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{forwards_noise2_Q.pdf}
        \caption{}
        \label{fig:egforwards2noiseQ}
    \end{subfigure}
    \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{forwards_noise2_R.pdf}
        \caption{}
        \label{fig:egforwards2noiseR}
    \end{subfigure}
    \caption{Top: Example of the simulated camera response spectra from each forward single-layer analytical model: Yudovsky 2009 (\textcolor{MyGreen}{green dotted}) and Jacques 1999 (\textcolor{MyOrange}{orange dashed}), using ground truth variables for a refractive index of 1.44 compared to that predicted by Monte Carlo with (\textcolor{purple}{purple dash-dotted} and without noise added (\textcolor{MyBlue}{blue solid}) for quantitative (left) and relative (right) data. Bottom: Demonstration of the variability of quality of prediction between the simulated camera responses from the Yudovsky 2009 two layer model (\textcolor{MyGreen}{green dotted}) and Monte Carlo with (\textcolor{purple}{purple dash-dotted} and without noise added (\textcolor{MyBlue}{blue solid}) when using the same ground truth parameters for quantitative (left) or relative (right) data.}
    \label{fig:forwardsHSIMC}
\end{figure}

@@

\begin{table}[h!]
    \centering
    \caption{The Pearson $r$ values (bold if $p<0.05$) of the linear regression line between the fitted tissue parameters and their ground truth displayed with their median absolute percentage errors ($APE$). This is shown for each variable when extracted by fitting Yudovsky 2009 single layer (Y1), Jacques 1999 (J), or Yudovsky 2009 double layer (Y2) to the quantitative (Q) or relative (R) Monte-Carlo datasets for $n=1.44$ at each noise level (0, 0.01, 0.03). All presented to 3s.f.}
    \begin{tabular}{|ccc|ccc|ccc|}
        \hline
        Parameter & Model & Quantitative & \multicolumn{3}{c}{$r$} & \multicolumn{3}{|c|}{median} \\
        & & (Q) or & \multicolumn{3}{c}{} & \multicolumn{3}{|c|}{$APE$ (\%)} \\
        & & Relative (R) & 0 & 0.01 & 0.03 & 0 & 0.01 & 0.03 \\
        \hline
        \multirow{6}{*}{$StO_2$} & \multirow{2}{*}{Y1} & Q & \textbf{0.999} & \textbf{0.825} & \textbf{0.369} & 1.90 & 24.8 & 85.2 \\
        & & R & \textbf{0.999} & \textbf{0.752} & \textbf{0.410} & 2.23 & 41.4 & 76.7 \\
        \cline{2-9}
        & \multirow{2}{*}{J} & Q & \textbf{0.978} & \textbf{0.829} & \textbf{0.402} & 4.06 & 22.5 & 83.6 \\
        & & R & \textbf{0.977} & \textbf{0.755} & \textbf{0.447} & 4.77 & 40.3 & 76.7 \\
        \cline{2-9}
        & \multirow{2}{*}{Y2} & Q & \textbf{0.751} & 0.0612 & 0.0194 & 16.8 & 66.5 & 68.9 \\
        & & R & \textbf{0.679} & \textbf{0.198} & 0.0448 & 18.8 & 63.0 & 75.1 \\
        \hline
        \multirow{6}{*}{$f_{blood}$} & \multirow{2}{*}{Y1} & Q & \textbf{0.984} & \textbf{0.667} & \textbf{0.450} & 5.83 & 23.7 & 51.8\\
        & & R & \textbf{0.870} & \textbf{0.547} & \textbf{0.535} & 16.7 & 47.5 & 51.1\\
        \cline{2-9}
        & \multirow{2}{*}{J} & Q & \textbf{0.924} & \textbf{0.719} & \textbf{0.498} & 7.14 & 25.0 & 46.7 \\
        & & R & \textbf{0.859} & \textbf{0.681} & \textbf{0.644} & 24.1 & 39.6 & 45.0 \\
        \cline{2-9}
        & \multirow{2}{*}{Y2} & Q & \textbf{0.872} & 0.147 & 0.0329 & 23.1 & 115 & 160 \\
        & & R & \textbf{0.609} & 0.189 & 0.0466 & 37.6 & 97.7 & 108 \\
        \hline
        \multirow{6}{*}{$a$} & \multirow{2}{*}{Y1} & Q & \textbf{0.992} & \textbf{0.773} & \textbf{0.558} & 3.85 & 18.7 & 35.1 \\
        & & R & \textbf{0.703} & 0.160 & \textbf{0.558} & 26.2 & 62.3 & 35.1 \\
        \cline{2-9}
        & \multirow{2}{*}{J} & Q & \textbf{0.961} & \textbf{0.759} & \textbf{0.542} & 4.69 & 17.1 & 30.4 \\
        & & R & \textbf{0.431} & -0.00356 & 0.166 & 27.7 & 57.7 & 51.7 \\
        \cline{2-9}
        & \multirow{2}{*}{Y2} & Q & \textbf{0.674} & 0.00365 & 0.0666 & 17.9 & 37.8 & 37.4 \\
        & & R & \textbf{0.333} & 0.0199 & -0.100 & 27.0 & 36.7 & 42.2 \\
        \hline
        \multirow{6}{*}{$b$} & \multirow{2}{*}{Y1} & Q & \textbf{0.998} & \textbf{0.751} & \textbf{0.516} & 2.63 & 21.8 & 52.0 \\
        & & R & \textbf{0.996} & \textbf{0.607} & \textbf{0.426} & 3.31 & 39.5 & 51.2 \\
        \cline{2-9}
        & \multirow{2}{*}{J} & Q & \textbf{0.934} & \textbf{0.757} & \textbf{0.515} & 5.70 & 23.7 & 50.4 \\
        & & R & \textbf{0.938} & \textbf{0.641} & \textbf{0.441} & 7.66 & 32.6 & 46.4 \\
        \cline{2-9}
        & \multirow{2}{*}{Y2} & Q & \textbf{0.735} & 0.169 & -0.00249 & 16.8 & 56.4 & 61.3 \\
        & & R & \textbf{0.360} & 0.0441 & -0.0711 & 33.2 & 61.7 & 63.6 \\
        \hline
        \multirow{2}{*}{$f_{mel}$} & \multirow{2}{*}{Y2} & Q & \textbf{0.852} & \textbf{0.622} & \textbf{0.477} & 28.0 & 85.1 & 90.1 \\ 
        & & R & \textbf{0.517} & 0.166 & \textbf{0.201} & 35.3 & 108 & 132 \\
        \hline
        \multirow{2}{*}{$L_1$} & \multirow{2}{*}{Y2} & Q & \textbf{0.824} & \textbf{0.401} & 0.123 & 36.7 & 69.5 & 84.5 \\
        & & R & \textbf{0.806} & 0.0777 & -0.0675 & 38.1 & 65.4 & 76.6 \\
        \hline
    \end{tabular}    
    \label{tb:backwardsHSIMC}
\end{table}

%could reduce number of things shown below/put some in supplementary but unsure
\begin{figure}[h!]
    \centering
    \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{backwardsn0Q.pdf}
        \caption{}
        \label{fig:backwardsn0Q}
    \end{subfigure}
    \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{backwardsn0R.pdf}
        \caption{}
        \label{fig:backwardsn0R}
    \end{subfigure}
    \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{backwardsn0.01Q.pdf}
        \caption{}
        \label{fig:backwardsn0.01Q}
    \end{subfigure}
    \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{backwardsn0.01R.pdf}
        \caption{}
        \label{fig:backwardsm0.01R}
    \end{subfigure}
    \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{backwardsn0.03Q.pdf}
        \caption{}
        \label{fig:backwardsn0.03Q}
    \end{subfigure}
    \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{backwardsn0.03R.pdf}
        \caption{}
        \label{fig:backwardsm0.03R}
    \end{subfigure}
    \caption{An example of the quality of parameter recovery by fitting each inverse analytical model (Yudovsky 2009 (\textcolor{MyGreen}{green $+$}) and Jacques 1999 (\textcolor{MyOrange}{orange $\times$})) to the quantitative (left) or relative (right) simulated camera responses from the Monte Carlo simulations and their associated trend lines for a refractive index of 1.44 for $StO_2$. This is shown for data with either no added noise (top) or added noise with $\sigma = 0.01$ (middle) or $\sigma = 0.03$ (bottom).}
    \label{fig:backwardsHSIMC1}
\end{figure}
